A Virtual Research Environment to Document and Analyze Non-alphabetic Writing Systems. A Case Study for Maya Writing

### 2016-02-10 source/31/EAGLE2016_I Sicily_final copy.tex
orfans and widows
percorsi tutti i file per eliminare orfane e vedove inserendo \newpage
o \vspace per spostare leggermente i paragrafi

\section{The motivations for and origins of \emph{I.Sicily}}
\noindent The existing epigraphic landscape in Sicily is extremely diverse in two primary regards: on the one hand, the island has
a very mixed cultural and linguistic make-up, meaning that the epigraphic material is itself extremely varied, with
extensive use throughout antiquity of both Greek and Latin, as well as Oscan, Punic, Sikel, and Hebrew;\footnote{
Recent overview of much of the linguistic tradition in \citet{tribulato_language_2012}; and of the epigraphic material in \citet{gulletta_sicilia_1999}.} on the other hand, the publication of this material has a very uneven record and despite an excellent
pre-twentieth-century tradition, the existing \emph{copora} are far from complete and the ability of key journals such as \emph{SEG}
or \emph{AE} to keep pace with local publication has been limited.\footnote{ For an overview of the \emph{corpora} tradition up to
the twentieth century, see \citet{de_vido_corpora_1999}.} A limited number of museum-based \emph{corpora} have been published in recent
decades (for Catania, Palermo, Messina, and Termini Imerese, as well as the material from Lipari), but this has not
greatly improved the overall situation.

The combination of these two factors already means that locating, identifying, or working with a Sicilian inscription,
or its publication record, is extremely challenging for anyone without extensive experience of the material. The
situation is compounded by the universal and familiar challenges of the recording and accessibility of archaeological
collections, whether held in museums, in archaeological stores, or elsewhere, and the lack of consistency in the
publication of \ new material.

@@

@@

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=0.6\columnwidth]{EAGLE2016ISicilyfinalcopy-img001.jpg}
\caption{sample edition}
\label{fig:1}
\end{figure}

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016ISicilyfinalcopy-img002.png}
\caption{Screenshot of map-based searching (museum locations) and of part of the spreadsheet/grid search interface
employed in \emph{I.Sicily}}
\label{fig:2}
\end{figure}

@@

\noindent The scale of the enterprise, and the available resources, mean that in its current form the project has limited itself
to inscriptions engraved on stone (the coverage of rupestral inscriptions/graffiti and of inscriptions painted on
stone/plaster is regrettably uneven). However, there is no reason in principle not to extend coverage in future to
include inscriptions on other materials. Similarly, although as noted above the current project does not include a
programme to mark up linguistic features of the texts, the commitment to the long-term maintenance of the corpus and
the open availability of the underlying XML records means that such a project would be entirely possible in the future.

A core principal of the project is that wherever possible an inscription record should be supported by recent autopsy
and not simply derived from the existing literature. Necessarily, this process is a slow one, and the majority of
records at this stage consist of information derived from secondary sources (earlier editions and other publications).
Individual inscription records will contain a clear indication of the editorial state of the record (from unchecked
through to fully edited) and additionally whether the record is underpinned by autopsy. In both cases, clear records
will be kept of editorial responsibility, autopsy and authorship as appropriate. In order to speed up the development
of the \emph{corpus}, and to encourage those working on the material to take ownership of it for themselves, we aim to enable
individuals to submit new records and emendations or additions to existing records (such as translations, images,
location information), both in the Epigraphic database and the Museums database. To this end, we welcome collaboration
with those undertaking epigraphic projects in Sicily, and aim to offer the ability for other projects to publish their
editions through \emph{I.Sicily}.\footnote{ We are currently establishing just such a collaboration with Prof.ssa M. Sgarlata
and Drs Lo Faro and Gradante, in support of their project to produce the volume of Inscriptiones Christianae Italiae
for the San Giovanni Catacombs in Siracusa.} We are also exploring the potential of the \emph{corpus} as a teaching resource
both for epigraphy in general and for the teaching of EpiDoc. This latter aspect has already been initiated through a
Teaching Project Award (2014-2015) from the Humanities Division of the University of Oxford, and we aim to develop this
further in the coming year, as part of the work of incorporation and conversion of texts into the existing dataset.

>>>>

\vspace{20pt}
\section{The motivations for and origins of \emph{I.Sicily}}
\noindent The existing epigraphic landscape in Sicily is extremely diverse in two primary regards: on the one hand, the island has
a very mixed cultural and linguistic make-up, meaning that the epigraphic material is itself extremely varied, with
extensive use throughout antiquity of both Greek and Latin, as well as Oscan, Punic, Sikel, and Hebrew;\footnote{
Recent overview of much of the linguistic tradition in \citet{tribulato_language_2012}; and of the epigraphic material in \citet{gulletta_sicilia_1999}.} on the other hand, the publication of this material has a very uneven record and despite an excellent
pre-twentieth-century tradition, the existing \emph{copora} are far from complete and the ability of key journals such as \emph{SEG}
or \emph{AE} to keep pace with local publication has been limited.\footnote{ For an overview of the \emph{corpora} tradition up to
the twentieth century, see \citet{de_vido_corpora_1999}.} A limited number of museum-based \emph{corpora} have been published in recent
decades (for Catania, Palermo, Messina, and Termini Imerese, as well as the material from Lipari), but this has not
greatly improved the overall situation.

\vspace{5pt}

The combination of these two factors already means that locating, identifying, or working with a Sicilian inscription,
or its publication record, is extremely challenging for anyone without extensive experience of the material. The
situation is compounded by the universal and familiar challenges of the recording and accessibility of archaeological
collections, whether held in museums, in archaeological stores, or elsewhere, and the lack of consistency in the
publication of new material.

\vspace{5pt}

@@

\vspace{5pt}

@@

\vspace{20pt}

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=0.6\columnwidth]{EAGLE2016ISicilyfinalcopy-img001.jpg}
\caption{Sample edition}
\label{fig:1}
\end{figure}

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016ISicilyfinalcopy-img002.png}
\caption{Screenshot of map-based searching (museum locations) and of part of the spreadsheet/grid search interface
employed in \emph{I.Sicily}}
\label{fig:2}
\end{figure}

@@

\noindent The scale of the enterprise, and the available resources, mean that in its current form the project has limited itself
to inscriptions engraved on stone (the coverage of rupestral inscriptions/graffiti and of inscriptions painted on
stone/plaster is regrettably uneven). However, there is no reason in principle not to extend coverage in future to
include inscriptions on other materials. Similarly, although as noted above the current project does not include a
programme to mark up linguistic features of the texts, the commitment to the long-term maintenance of the corpus and
the open availability of the underlying XML records means that such a project would be entirely possible in the future.
\newpage
A core principal of the project is that wherever possible an inscription record should be supported by recent autopsy
and not simply derived from the existing literature. Necessarily, this process is a slow one, and the majority of
records at this stage consist of information derived from secondary sources (earlier editions and other publications).
Individual inscription records will contain a clear indication of the editorial state of the record (from unchecked
through to fully edited) and additionally whether the record is underpinned by autopsy. In both cases, clear records
will be kept of editorial responsibility, autopsy and authorship as appropriate. In order to speed up the development
of the \emph{corpus}, and to encourage those working on the material to take ownership of it for themselves, we aim to enable
individuals to submit new records and emendations or additions to existing records (such as translations, images,
location information), both in the Epigraphic database and the Museums database. To this end, we welcome collaboration
with those undertaking epigraphic projects in Sicily, and aim to offer the ability for other projects to publish their
editions through \emph{I.Sicily}.\footnote{ We are currently establishing just such a collaboration with Prof.ssa M. Sgarlata
and Drs Lo Faro and Gradante, in support of their project to produce the volume of Inscriptiones Christianae Italiae
for the San Giovanni Catacombs in Siracusa.} We are also exploring the potential of the \emph{corpus} as a teaching resource
both for epigraphy in general and for the teaching of EpiDoc. This latter aspect has already been initiated through a
Teaching Project Award (2014-2015) from the Humanities Division of the University of Oxford, and we aim to develop this
further in the coming year, as part of the work of incorporation and conversion of texts into the existing dataset.

### 2016-02-10 source/38/EAGLE2016_Final_Paper Modolo.tex
Paper 38 (Modolo): long url in n.6 fixed
orfans and widows
percorsi tutti i file per eliminare orfane e vedove inserendo \newpage
o \vspace per spostare leggermente i paragrafi

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{Free Reproduction of Cultural Heritage for Research Purposes:
Regulatory Aspects and New Prospects for Project EAGLE}
\author[free]{Mirco Modolo\corref{first}}
\address[free]{Free lance archaeologist and archivist}
\cortext[first]{Corresponding author. Email: mircomodolo@gmail.com}
\date{2015-11-20}
\begin{document}

@@

Moreover, the movement produced a proposal to amend article 108 of the Codice dei Beni Culturali regulating the
reproduction of cultural heritage,\footnote{\url{https://fotoliberebbcc.wordpress.com/category/la-nostra-proposta/}} to eliminate the
exclusion of printed books and documents from the liberalisation archive, hence complying with copyright and personal
privacy for archival documents.

Free photography is therefore a fundamental tool which, without contradicting the demands of conservation, greatly
facilitates the task of transcribing documents and historical research – allowing significant time and money savings
for the scholar, especially those forced to move to archives distant from their own place of residence.

>>>>

\usepackage{url}
\def\UrlBreaks{\do\/\do-}
%%%%%%%%
%TITLE:%
%%%%%%%%
\title{Free Reproduction of Cultural Heritage for Research Purposes:
Regulatory Aspects and New Prospects for Project EAGLE}
\author[free]{Mirco Modolo\corref{first}}
\address[free]{Free lance archaeologist and archivist}
\cortext[first]{Corresponding author. Email: mircomodolo@gmail.com}
\date{2015-11-20}
\begin{document}

@@

Moreover, the movement produced a proposal to amend article 108 of the Codice dei Beni Culturali regulating the
reproduction of cultural heritage,\footnote{\url{https://fotoliberebbcc.wordpress.com/category/la-nostra-proposta/}} to eliminate the
exclusion of printed books and documents from the liberalisation archive, hence complying with copyright and personal
privacy for archival documents.
\newpage
Free photography is therefore a fundamental tool which, without contradicting the demands of conservation, greatly
facilitates the task of transcribing documents and historical research – allowing significant time and money savings
for the scholar, especially those forced to move to archives distant from their own place of residence.

### 2016-02-10 source/42/Paper_licciardello.tex
orfans and widows
percorsi tutti i file per eliminare orfane e vedove inserendo \newpage
o \vspace per spostare leggermente i paragrafi

My research will try to detect different trends in the use of deictic expressions in order to highlight the similarities
and parallelisms between inscribed and literary epigrams and to elucidate the different deictic strategies employed in
different contexts.

@@

The corpus of epigrams analysed includes dedicatory verse-inscriptions collected in Hansen's Carmina Epigraphica Graeca
[CEG 1-2], as for epigrams on stone. In this case, I selected those epigrams included by Hansen in the section ``Tituli
dedicatorii''. For literary epigrams, I followed Gow's and Page's edition \citep{gow_greek_1965}, to which I added
the epigrams attributed to Posidippus and edited by Austin and Bastianini \citep{austin_posidippi_2002}. In this case, the
selection was made according to the contents of the epigrams and I selected all those epigrams that have a clear
dedicatory frame.

The structure of the database was designed to include all important data related to the text examined, in order to have
for each epigram a complete profile, which includes linguistic, literary, historical and archaeological
aspects.\footnote{ The software used is FileMaker Pro 13.0v4.} The data were organised in various tables, which are all interconnected (See Fig.~\ref{fig:1}).

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{schema.png}
\caption{default}
\label{fig:1}
\end{figure}

>>>>

\vspace{5pt}
My research will try to detect different trends in the use of deictic expressions in order to highlight the similarities
and parallelisms between inscribed and literary epigrams and to elucidate the different deictic strategies employed in
different contexts.

@@

\vspace{5pt}
The corpus of epigrams analysed includes dedicatory verse-inscriptions collected in Hansen's Carmina Epigraphica Graeca
[CEG 1-2], as for epigrams on stone. In this case, I selected those epigrams included by Hansen in the section ``Tituli
dedicatorii''. For literary epigrams, I followed Gow's and Page's edition \citep{gow_greek_1965}, to which I added
the epigrams attributed to Posidippus and edited by Austin and Bastianini \citep{austin_posidippi_2002}. In this case, the
selection was made according to the contents of the epigrams and I selected all those epigrams that have a clear
dedicatory frame.

\vspace{5pt}
The structure of the database was designed to include all important data related to the text examined, in order to have
for each epigram a complete profile, which includes linguistic, literary, historical and archaeological
aspects.\footnote{ The software used is FileMaker Pro 13.0v4.} The data were organised in various tables, which are all interconnected (See Fig.~\ref{fig:1}).

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{schema.png}
\caption{default}
\label{fig:1}
\end{figure}

### 2016-02-10 source/43/EAGLE2016_Benefiel_Sypniewski.tex
corrections made to 43, 33, 29
orfans and widows
percorsi tutti i file per eliminare orfane e vedove inserendo \newpage
o \vspace per spostare leggermente i paragrafi

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BenefielSypniewski-img005.jpg}
\caption{Example of an entry for a figural graffito not found in \emph{CIL} (EDR144514)}
\label{fig:5}
\end{figure}

@@

As one might imagine, issues of interpretation can arise even with simple descriptions of images. In truth, we have
encountered more difficulties with interpretation in the case of drawings that have been described by \emph{CIL}. The first
example comes from a shop in Pompeii and represents text and a drawing [\emph{CIL} 4, 8185]. \ The plaster has clearly broken
off, so we do not know if this was part of a larger scene. What remains are two lines of text and just one figure,
which would seem to be a drawing of a person facing forward and rendered with head and shoulders. \emph{CIL} describes it
thus: \emph{herma \underline{muliebris} prospiciens} [Fig.~\ref{fig:6}].

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BenefielSypniewski-img006.jpg}
\caption{\emph{CIL} entry for textual and figural graffiti [\emph{CIL} 4, 8185]}
\label{fig:6}
\end{figure}

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BenefielSypniewski-img007.jpg}
\caption{Apograph of figural graffito referred to in the note at \emph{CIL} 4, 10676 \citet[n. 309]{langner_antike_2001}.}
\label{fig:7}
\end{figure}

@@

@@

\noindent The third challenge that we face is how to search efficiently for inscriptions that are not just text but either are
images or include images. In the AGP search engine, we aim to complement the capabilities of EDR by providing another
way to search for these non-textual, figural graffiti. Since we describe the content of the figural graffito in the
Textus field of EDR, it is possible for a user to locate a graffito drawing. However, with text-based searching, a user
would need to need to know the vocabulary used to describe the drawing. Would someone ever think to search for
``\emph{camelus}'' without prior knowledge that there is a figural graffito of a camel in Herculaneum? Similarly, if you
search for ``\emph{gladiator}'', the text field will give you results for all inscriptions that mention gladiators as well as
drawings where we have described gladiators. If, however, we’ve described the gladiator more specifically as a
``\emph{retiarius}'' or we have gladiatorial equipment, such drawings will be omitted from the list of search results.

We are therefore designing AGP with the capacity for locating figural graffiti through a two-prong solution: with
both browsing and searching possibilities.

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BenefielSypniewski-img008.png}
\caption{Screenshot of AGP search engine showing browsing capabilities.}
\label{fig:8}
\end{figure}

@@

\subsection{Limiting search results in AGP}

\noindent It is therefore necessary to design a way to limit results, so that a user could find only gladiatorial equipment,
or graffiti depicting only boars but not graffiti depicting other animals. We are developing a method of filters and
tags that will allow a user to move beyond browsing. These filters will allow a user 1) to limit an initial return of
results, 2) to retrieve more specific results, or 3) to perform a secondary level of search. It will certainly be
helpful to refine results of an entire category to include only a subset of that category, for example, only pairs of
gladiators instead of all gladiators and their equipment.

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BenefielSypniewski-img009.png}
\caption{Screenshot showing filters in AGP.}
\label{fig:9}
\end{figure}

Or, it will be possible to do a broad search for all drawings of animals, and then filter to limit the results to find
what kind of animals are drawn in taverns, for example, but not public buildings or houses. \

>>>>

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BenefielSypniewski-img005.jpg}
\caption{Example of an entry for a figural graffito not found in \emph{CIL} (EDR144514)}
\label{fig:5}
\end{figure}

@@

As one might imagine, issues of interpretation can arise even with simple descriptions of images. In truth, we have
encountered more difficulties with interpretation in the case of drawings that have been described by \emph{CIL}. The first
example comes from a shop in Pompeii and represents text and a drawing [\emph{CIL} 4, 8185]. \ The plaster has clearly broken
off, so we do not know if this was part of a larger scene. What remains are two lines of text and just one figure,
which would seem to be a drawing of a person facing forward and rendered with head and shoulders. \emph{CIL} describes it
thus: \emph{herma \underline{muliebris} prospiciens} [Fig.~\ref{fig:6}].
\newpage

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BenefielSypniewski-img006.jpg}
\caption{\emph{CIL} entry for textual and figural graffiti [\emph{CIL} 4, 8185]}
\label{fig:6}
\end{figure}

@@

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BenefielSypniewski-img007.jpg}
\caption{Apograph of figural graffito referred to in the note at \emph{CIL} 4, 10676 \citet[n. 309]{langner_antike_2001}.}
\label{fig:7}
\end{figure}

@@

\noindent The third challenge that we face is how to search efficiently for inscriptions that are not just text but either are
images or include images. In the AGP search engine, we aim to complement the capabilities of EDR by providing another
way to search for these non-textual, figural graffiti. Since we describe the content of the figural graffito in the
Textus field of EDR, it is possible for a user to locate a graffito drawing. However, with text-based searching, a user
would need to need to know the vocabulary used to describe the drawing. Would someone ever think to search for
``\emph{camelus}'' without prior knowledge that there is a figural graffito of a camel in Herculaneum? Similarly, if you
search for ``\emph{gladiator}'', the text field will give you results for all inscriptions that mention gladiators as well as
drawings where we have described gladiators. If, however, we’ve described the gladiator more specifically as a
``\emph{retiarius}'' or we have gladiatorial equipment, such drawings will be omitted from the list of search results.
\newpage
We are therefore designing AGP with the capacity for locating figural graffiti through a two-prong solution: with
both browsing and searching possibilities.

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=0.9\columnwidth]{EAGLE2016BenefielSypniewski-img008.png}
\caption{Screenshot of AGP search engine showing browsing capabilities.}
\label{fig:8}
\end{figure}

@@

\subsection{Limiting search results in AGP}
\noindent It is therefore necessary to design a way to limit results, so that a user could find only gladiatorial equipment,
or graffiti depicting only boars but not graffiti depicting other animals. We are developing a method of filters and
tags that will allow a user to move beyond browsing. These filters will allow a user 1) to limit an initial return of
results, 2) to retrieve more specific results, or 3) to perform a secondary level of search. It will certainly be
helpful to refine results of an entire category to include only a subset of that category, for example, only pairs of
gladiators instead of all gladiators and their equipment.

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=0.9\columnwidth]{EAGLE2016BenefielSypniewski-img009.png}
\caption{Screenshot showing filters in AGP.}
\label{fig:9}
\end{figure}

Or, it will be possible to do a broad search for all drawings of animals, and then filter to limit the results to find
what kind of animals are drawn in taverns, for example, but not public buildings or houses.

### 2016-02-10 source/48/EAGLE16-lame-et-al-teaching.tex
orfans and widows
percorsi tutti i file per eliminare orfane e vedove inserendo \newpage
o \vspace per spostare leggermente i paragrafi

 A traditional print photograph, such as Fig.~\ref{fig:12582} (No. 12582), drawn from the archives of the AMZ, limits the student’s
opportunity to learn to decipher these tags. Even if a professional photographer took this picture with a good quality
camera in a fully equipped photographic studio, the reading of the inscription stays riddled with uncertainties and a
researcher would have to check the tag \emph{de visu} in order to offer a plausible transcription.

@@

Traces of older inscriptions are hard to read and interpret but several personal names may be surmised. Such a tag was
particularly interesting for our experiment. It is quite a challenging inscription, even for a skilled epigraphist and
paleographer; we therefore were anxious to find out how efficiently the students would tackle it. Thanks to the DAP,
the transcription was not utterly difficult for most students. Obviously, in the first instance, the interpretation was
beyond their means, but we were nevertheless impressed by their abilities to read it far more easily than expected. The
DAP definitely offers the possibility to work on such material online, without the obligation to inspect it personally,
at least in the first stage.

@@

The older, erased layers, on the obverse of tag No. 13053, drawn with dashed lies on the drawings of the academic edition, are more difficult to read. No
g/t relationship between wSystem and tSystem was asked to students (see Fig.~\ref{fig:4}).

@@

\section*{Aknowledgment}
\noindent We wish to acknowledge the research assistance of T. M. Crandall,
History student at Mount Allison University (Sackville, Canada), and C. Le Borgne, a Roman epigrapher and student of
M. Smith, who is Professor of Paleography at the École Nationale des Chartes and at the École Pratique des Hautes
Études (PSL Research University). His current seminar, for the Academic Year 2015/16, at the EPHE is entitled
``La lettre romaine, entre inscriptions monumentales et écritures communes''. The following
students at Mount Allison University participated in this experiment as part of their course work: D. Atkinson, E.
Beaton, M. Gillard, T. Hall and K. Wallace. We warmly thank M. Vazzana and C. Tanca, students of the Laurea
Specialistica in Informatica Umanistica at the University of Pisa, who prepared part of the great pedagogical material
as well as some useful didactic videos for the Laboratorio di Cultura Digitale. We are grateful to the Deutsches
Archäologisches Institut, and especially to Prof. R. Förtsch and to M. Romanello for offering us the opportunity to
present the DAP to the students of the Digital Classicist Berlin Seminar the 8th of December 2015.
\nocite{Lame2015b}
\nocite{cherry_storytelling_2014}
\bibliographystyle{sapauth-eng}
\bibliography{../../EAGLE}

>>>>

\newpage
 A traditional print photograph, such as Fig.~\ref{fig:12582} (No. 12582), drawn from the archives of the AMZ, limits the student’s
opportunity to learn to decipher these tags. Even if a professional photographer took this picture with a good quality
camera in a fully equipped photographic studio, the reading of the inscription stays riddled with uncertainties and a
researcher would have to check the tag \emph{de visu} in order to offer a plausible transcription.

@@

Traces of older inscriptions are hard to read and interpret but several personal names may be surmised. Such a tag was
particularly interesting for our experiment. It is quite a challenging inscription, even for a skilled epigraphist and
paleographer; we therefore were anxious to find out how efficiently the students would tackle it. Thanks to the DAP,
the transcription was not utterly difficult for most students. Obviously, in the first instance, the interpretation was
beyond their means, but we were nevertheless impressed by their abilities to read it far more easily than expected.

\newpage
The DAP definitely offers the possibility to work on such material online, without the obligation to inspect it personally,
at least in the first stage.

@@

\newpage
The older, erased layers, on the obverse of tag No. 13053, drawn with dashed lies on the drawings of the academic edition, are more difficult to read. No
g/t relationship between wSystem and tSystem was asked to students (see Fig.~\ref{fig:4}).

@@

\vspace{10pt}
\section*{Aknowledgment}
\noindent We wish to acknowledge the research assistance of T. M. Crandall,
History student at Mount Allison University (Sackville, Canada), and C. Le Borgne, a Roman epigrapher and student of
M. Smith, who is Professor of Paleography at the École Nationale des Chartes and at the École Pratique des Hautes
Études (PSL Research University). His current seminar, for the Academic Year 2015/16, at the EPHE is entitled
``La lettre romaine, entre inscriptions monumentales et écritures communes''. The following
students at Mount Allison University participated in this experiment as part of their course work: D. Atkinson, E.
Beaton, M. Gillard, T. Hall and K. Wallace. We warmly thank M. Vazzana and C. Tanca, students of the Laurea
Specialistica in Informatica Umanistica at the University of Pisa, who prepared part of the great pedagogical material
as well as some useful didactic videos for the Laboratorio di Cultura Digitale. We are grateful to the Deutsches
Archäologisches Institut, and especially to Prof. R. Förtsch and to M. Romanello for offering us the opportunity to
present the DAP to the students of the Digital Classicist Berlin Seminar the 8th of December 2015.
\nocite{Lame2015b}
\nocite{cherry_storytelling_2014}
\bibliographystyle{sapauth-eng}
\bibliography{../../EAGLE}

### 2016-02-10 source/8/EAGLE_2016_submission_8_revised.tex
orfans and widows
percorsi tutti i file per eliminare orfane e vedove inserendo \newpage
o \vspace per spostare leggermente i paragrafi

The form of a text carrier and its function also influence the text. Different lexemes may thus be determined for
describing round or quadrilateral altars, with the result that conclusions may be drawn concerning the semantics and
possible linguistic decipherments. Physical description of the text carrier as an object and its linguistic,
historical, and cultural contextualization are indispensable for deciphering a previously unknown text or analyzing its
arrangement. These metadata must be consulted in order to grammatologically and linguistically analyze the text’s
contents.

Modelling an object biography and culturally contextualizing the texts must be initially addressed with information
technology in order to facilitate epigraphic analysis of the text’s contents in its extra-linguistic, cultural context.
Thus, it is necessary to create an object metadata schema that captures such information concerning the language’s
cultural area and makes it available for use.

@@

Annotating texts composed in such writing systems that use only alphabetic transcription is insufficient; instead, the
original spelling of a lemma should be represented. This approach facilitates studies of paleography or sign usage
across time and space. In addition, it can reveal preferred sign arrangements within a block or preferences for
particular signs that correspond to the material used or to the subject of the text, for instance.

Thus, when creating a standard for non-alphabetic scripts, object metadata and annotations for orthographic and
linguistic analysis must be taken into account, and the creative process should also promote discussion of
terminological and typographic conventions for these annotations \citep{SachseDurr2015}. The goal is to create a generic
markup model that can be implemented regardless of the particular script or language in question. The inscriptions of
the Maya hieroglyphic writing system alone feature attestations of various vernacular languages \citep{LacadenaWichmann2002}, a phenomenon which raises several preliminary questions: what exactly are the demands, and what are the goals of
an epigraphic and linguistic annotation? Where do possibilities and limitations exist for annotating these writing
systems using established standards, such as TEI or EpiDoc? Existing standards have to be modified to some extend and
possibly new standards have to be created, to overcome these limitations.

@@

\section[Summary ]{Summary}
\noindent The creation of digital infrastructures and metadata schemas that facilitate recording and further analysis of the
appropriate annotation for non-alphabetic writing systems, including their respective cultural contexts, is a
desideratum in the Digital Humanities that needs to be addressed.

>>>>

The form of a text carrier and its function also influence the text. Different lexemes may thus be determined for
describing round or quadrilateral altars, with the result that conclusions may be drawn concerning the semantics and
possible linguistic decipherments. Physical description of the text carrier as an object and its linguistic,
historical, and cultural contextualization are indispensable for deciphering a previously unknown text or analyzing its
arrangement. These metadata must be consulted in order to grammatologically and linguistically analyze the text’s
contents. 
\newpage
Modelling an object biography and culturally contextualizing the texts must be initially addressed with information
technology in order to facilitate epigraphic analysis of the text’s contents in its extra-linguistic, cultural context.
Thus, it is necessary to create an object metadata schema that captures such information concerning the language’s
cultural area and makes it available for use.

@@

Annotating texts composed in such writing systems that use only alphabetic transcription is insufficient; instead, the
original spelling of a lemma should be represented. This approach facilitates studies of paleography or sign usage
across time and space. In addition, it can reveal preferred sign arrangements within a block or preferences for
particular signs that correspond to the material used or to the subject of the text, for instance.
Thus, when creating a standard for non-alphabetic scripts, object metadata and annotations for orthographic and
linguistic analysis must be taken into account, and the creative process should also promote discussion of
terminological and typographic conventions for these annotations \citep{SachseDurr2015}. The goal is to create a generic
markup model that can be implemented regardless of the particular script or language in question. The inscriptions of
the Maya hieroglyphic writing system alone feature attestations of various vernacular languages \citep{LacadenaWichmann2002}, a phenomenon which raises several preliminary questions: what exactly are the demands, and what are the goals of
an epigraphic and linguistic annotation? Where do possibilities and limitations exist for annotating these writing
systems using established standards, such as TEI or EpiDoc? Existing standards have to be modified to some extend and
possibly new standards have to be created, to overcome these limitations.

@@

\newpage
\section[Summary ]{Summary}
\noindent The creation of digital infrastructures and metadata schemas that facilitate recording and further analysis of the
appropriate annotation for non-alphabetic writing systems, including their respective cultural contexts, is a
desideratum in the Digital Humanities that needs to be addressed.
